Why does preregistration increase the persuasiveness of evidence? A Bayesian rationalization.

### 2022-06-08 manuscript.Rmd

A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
These skeptics question why a study should be preregistered at all.
Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include, in how much detail in the preregistration, or about the conditions that warrant deviating from it.<!-- Message unclear, if meant like I think: "...such as what to include in the preregistraion and how detailed this should be"-->
If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.

@@

The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
However, if the authors offer a convincing argument for their deviation, their readers might be compelled to increase their uncertainty only a limited amount.
<!-- Last sentence is a bit confusing -->

>>>>

A successful prediction should lend more credibility to a theory than a postdiction, all else being equal.
The scientific community has long pondered the vital distinction between exploration and confirmation, discovery and justification, hypothesis-generating and hypothesis-testing and so forth [see @hoyningen-hueneContextDiscoveryContext2006].
Confusing exploratory findings with confirmed theories has lead to a crisis of confidence in the results of empirical sciences [@ioannidisWhyMostPublished2005] and psychology in particular [@opensciencecollaborationEstimatingReproducibilityPsychological2015].
As a response, more and more researchers preregister their studies [@nosekPreregistrationRevolution2018] to honour this distinction and produce results that are considered confirmatory.
Indeed, rigorous application of preregistration prevents researchers from reporting a simple confirmatory story for a set of results produced by an arduous process of trial and error [@wagenmakersAgendaPurelyConfirmatory2012].
However, while many researchers intuitively recognize this as an advantage of preregistration, some remain unconvinced.
These skeptics question why a study should be preregistered at all.
Even researchers convinced of the advantages of preregistration face difficult problems when they exclusively rely on an intuitive appeal.
Specifically, an intuitive conviction is unable to address pragmatic questions, such as what to include in the preregistraion and how detailed this should be, or about the conditions that warrant deviating from it.
If the research community remains confused about the purpose of preregistration, it is bound to misapply it and hence decrease the efficiency of the scientific endeavour. 
We thus propose a principled justification for preregistration with a formal objective against which researchers can evaluate several pragmatic trade-offs.

@@

The advantage of this perspective that puts uncertainty about theoretical risk front and center is twofold.
First, aiming to reduce uncertainty is beneficial across a wide range of potential theoretical risks.
That is, researchers benefit from preregistrating their study, regardless whether or not the study is "exploratory".
If researchers conduct a more exploratory study they can clearly communicate how exploratory they aim to be and their results can be judged accordingly.
Second, if researchers realize after the fact, that it would be unwise to follow their preregistration in certain details, they may change it.
Allowing to deviate from the preregistration is controversial (XXX) because the authors might sift through all the possibilities and then select the most beneficial for their theory.
This argument, in its fullest consequence, might result in the same uncertainty we argued is appropriate for not preregistered studies.
However, if the authors offer a convincing argument for their deviation, their readers might believe it to be a likely explanation for the deviation.
If the readers find the reason for the deviation likely to be something other than decreasing the theorethical risk, their uncertainty about the theorethical risk should only increase modestly.
